Iconic representations of molluscs through the ages

### 2012-10-25 large_number_squeezing_with_atoms.tex
Begin fleshing out the appendix

\section{Model and single mode solutions}
\begin{itemize}
  \item Lay out the scheme with diagram, define all variable etc
  \item Present multimode analytic solution, details go into an appendix
  \item Plot a few solutions to give an idea of what things look like, plausible times and so on
  \item Maybe point out it agrees with Oberthaler, or save that for discussion on MM effects?
  \item Point out that number *difference* squeezing can often be considerably higher than absolute number squeezing in a single state
\end{itemize}

@@

We allow the atoms to interact via a standard Kerr-type nonlinearity arising from atom-atom interactions, as well as allowing a linear coupling between the two modes. Denoting the two modes by $|a\rangle$ and $|b\rangle$, the Hamiltonian governing this system is given by
\begin{equation}
\hat{H} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b} 
          + \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
          + \chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
          + \frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}
          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{a}^{\dagger} \hat{b} )
\label{eqTwoModeHamiltonian}
\end{equation}
where the $\chi_{ij}$ describe the various inter- and intra-mode nonlinearities, and $\Omega$ couples the two modes and allows for population transfer between them.

@@

\appendix
\section{Derivation of single mode squeezing formula}
\label{appendixTwoModeDerivation}

>>>>

\section{Model and two mode solutions}
\label{secTwoModeAnalytic}
\begin{itemize}
  \item Lay out the scheme with diagram, define all variable etc
  \item Present multimode analytic solution, details go into an appendix
  \item Plot a few solutions to give an idea of what things look like, plausible times and so on
  \item Maybe point out it agrees with Oberthaler, or save that for discussion on MM effects?
  \item Point out that number *difference* squeezing can often be considerably higher than absolute number squeezing in a single state
\end{itemize}

@@

We allow the atoms to interact via a standard Kerr-type nonlinearity arising from atom-atom interactions, as well as allowing a linear coupling between the two modes. Denoting the two modes by $|a\rangle$ and $|b\rangle$, the Hamiltonian governing this system is given by
\begin{equation}
\hat{H} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b} 
          + \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
          + \chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
          + \frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}
          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger}  \hat{a} )
\label{eqTwoModeHamiltonian}
\end{equation}
where the $\chi_{ij}$ describe the various inter- and intra-mode nonlinearities, and $\Omega$ couples the two modes and allows for population transfer between them.

@@

\appendix
\section{Derivation of two mode squeezing formula}
\label{appendixTwoModeDerivation}
In this Appendix we derive the equations for the number variances which were earlier stated without proof. We use the model described in Section \ref{secTwoModeAnalytic}, where we consider a two mode system with $N$ particles, initially prepared as a coherent state with all the population in mode $|a\rangle$ at time $t=0$. A coupling pulse is then applied, so that the system evolves under the Hamiltonian
\begin{equation}
\hat{H_1} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b} 
          + \Omega (\hat{a}^{\dagger} \hat{b} + \hat{b}^{\dagger} \hat{a} )
\end{equation}
until time $t=t_1$. From time $t_1$ until time $t=t_2$ the coupling is turned off, and the system evolves under the Hamiltonian
\begin{equation}
\hat{H_2} = \frac{\chi_{aa}}{2} \hat{a}^{\dagger} \hat{a}^{\dagger} \hat{a} \hat{a}
          + \chi_{ab} \hat{a}^{\dagger} \hat{a} \hat{b}^{\dagger} \hat{b}
          + \frac{\chi_{bb}}{2} \hat{b}^{\dagger} \hat{b}^{\dagger} \hat{b} \hat{b}.
\end{equation}
Finally, from time $t_2$ until time $t=t_3$, the coupling is once again applied with a phase offset of $\phi$ relative to the first time, and the system evolves via
\begin{equation}
\hat{H_3} = \hbar \omega \hat{a}^{\dagger} \hat{a} +  \hbar \omega \hat{b}^{\dagger} \hat{b}
          + \Omega (e^{i\phi} \hat{a}^{\dagger} \hat{b} + e^{-i\phi} \hat{b}^{\dagger} \hat{a} )
\end{equation}
It is clear that the linear self-energy terms proportional to $\omega$ result in a constant phase rotation of the fields $\hat{a}$ and $\hat{b}$. As this system is self-contained and global phase is unmeasurable, all phase measurements can only rely on the difference in phase between the two modes. This phase rotation is therefore irrelevent and will be ignored in the following analysis.

We will work in the Heisenberg picture, and derive expressions for $\hat{a}(t_3)$ and $\hat{b}(t_3)$ in terms of $\hat{a}(t_2)$ and $\hat{b}(t_2)$, which in turn can be expressed in terms of $\hat{a}(t_1)$ and $\hat{b}(t_1)$. We will show that the state of the system at time $t_1$ consists of seperate coherent states in modes $|a\rangle$ and $|b\rangle$, enabling us to calculate expectation values at time $t_3$ in terms of the known state at $t_1$.

In order to have more compact expressions we will use the notation
\begin{eqnarray}
\hat{a}(t_j) &=& \hat{a}_j \\
\hat{b}(t_j) &=& \hat{b}_j.
\end{eqnarray}

Beginning with the Hamiltonian $H_3$, during the period $(t_2, t_3)$ mode $|a\rangle$ will evolve as
\begin{equation}
\hat{a}_3 = e^{ i H_3 t/ \hbar} \, \hat{a}_2 \, e^{-i H_3 t/ \hbar}.
\end{equation}
Utilizing the Hadamard lemma
\begin{equation}
e^X Y e^{-X} = Y + [X,Y] + \frac{1}{2!}[X,[X,Y]] + \frac{1}{3!}[X,[X,[X,Y]]] + \ldots
\end{equation}
we obtain
\begin{eqnarray}
\hat{a}_3 &=& \hat{a}_2 - i e^{i\phi} \Omega \tau \hat{b}_2 + \frac{(i \Omega \tau)^2}{2!} \hat{a}_2 - \frac{(i \Omega \tau)^3}{3!}e^{i\phi} \hat{b}_2 + \frac{(i \Omega \tau)^4}{4!} \hat{a}_2 - \ldots \nonumber \\
          &=& \cos (\Omega \tau) \hat{a}_2 -i e^{i\phi} \sin (\Omega\tau) \hat{b}_2
\end{eqnarray}
where $\tau=t_3-t_2$ is the period for which the coupling is applied. Similarly, we find
\begin{equation}
\hat{b}_3 = \cos (\Omega \tau) \hat{b}_2 -i e^{-i\phi} \sin (\Omega\tau) \hat{a}_2.
\end{equation}

(1) Find $\hat{a}_2$, $\hat{a}_2$ in terms of $\hat{a}_1$, $\hat{b}_1$.

(2) Derive 1631.5 and explain we need expectation values of the bilinear, trilinear and quartic terms. We can use result of previous step to do them, provided we normally order everything

(3) Take each set of terms from 1631.5 individually, normal order them, and evaluate

(4) Plug list of expectation values back into 1631.5
